---
title: "Normal Distribution in R"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
library(mosaic)
```


## Plotting normal distributions

In `R` we can plot a normal distribution curve with given mean and standard deviation using the `gf_dist` function:

For example, the following will plot a normal distribution with mean $\mu = 34$ and standard deviation $\sigma = 5$:

```{r, echo=TRUE}
gf_dist("norm", mean = 34, sd = 5)
```


### Try it yourself

Enter the code required to plot a normal distribution curve. Try to experiment with different means and standard deviations:

```{r normal-curve, exercise=TRUE}

```

### Shading areas

You can "shade" the areas under the curve, to visualize the probabilities more clearly by using `pdist` instead of `gf_dist`.  To split the areas under the normal curve into three parts: one to the left of -2, one between -2 as 1.5, and one to the right of 1.5, we do this (feel free to experiment with the code, and change the numbers.What would happen if you have only one number inside the `c(...)`?):

```{r normal-shaded, exercise=TRUE, exercise.eval=TRUE}
pdist("norm", mean = 0, sd = 1, c(-2, 1.5), return = "plot")
```

## Calculating probabilities

### Area to the left side

Let's look at the problem 6 in Chapter 10 of the textbook:

According to a study cited in the textbook,
the babies born in singleton births in the United States have birth weights that are approximately normally distributed with mean 3.296 kg and standard deviation 0.560 kg.

_What is the probability that a baby is born weighing less than 2.57 kg?_

Since the baby weight are normally distributed, we can use normal distribution to answer this.  For the probability of a baby weighing __less__ than 2.57 kg, we need to find the area to the __left__ of 2.57 under the normal curve with mean 3.296 and standard deviation 0.560.

For that, we use `pnorm`:

```{r, echo=TRUE}
pnorm(2.57, mean = 3.296, sd = 0.56)
```

Use `pnorm` to find the probability that a baby is born weighing less than 3.34 kg:

```{r small_baby, exercise=TRUE}

```

```{r small_baby-solution}
pnorm(3.34, mean = 3.296, sd = 0.56)
```

Use `pnorm` to find the area to the left of 34 under the normal curve with mean 31 and standard deviation 10:

```{r left_area, exercise=TRUE}

```

```{r left_area-solution}
pnorm(34, mean = 31, sd = 10)
```


### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```
